---
title: "Inferring Cow Phenotypes Based on Sensor Data"
author: "Peter von Rohr and Jessica Gearning"
date: '2021-06-18'
output:
  powerpoint_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Content

1. What are the data 
2. What can we do with this data
3. Introduction to Machine Learning
4. Everyday use cases of machine learning
5. Machine learning in agriculture
6. Second level algorithm
7. Conclusions


## Data

* Output of first level classification
* For every cow over time the basic behavior
* Basic behavior can be one of 
    + Eating
    + Laying
    + Standing
    + Walking
    + Ruminating


## Basic Behavior over Time

```{r tile-plot, echo=FALSE, message=FALSE, warning=FALSE}
n_nr_obs <- 30
vec_behave <- c("eating", "laying", "ruminating", "standing", "walking")
n_nr_behave <- length(vec_behave)
tbl_behavior <- tibble::tibble(x = 1:n_nr_obs, y = sample(vec_behave, n_nr_obs, replace = TRUE))
require(ggplot2)
p <- ggplot(tbl_behavior, aes(x=x, y=as.factor(y), fill = as.factor(y))) + xlab(NULL) + ylab(NULL)
p + geom_tile()
```


## Data Aggregation Over Time

* Summarize duration of behaviors over one day
* Yields Average time for a certain behavior per day
* Defines graphical representation of behavior pattern diagram

<!-- Insert geom_line plot of behavior, similarly to Alibaba or Bar-plot -->


## Overlay Phenotypes of Interest

* Phenotypes of interest
    + Heat
    + Lameness
    + Calving
    + Mastitis
    + Ketosis
    + ...
* Pooled as events where farmer must be alerted
* Overlay events in behavior pattern diagrams
* Associate events with behavior pattern diagrams


## Aim: Prediction of Events

* Given association of events and cow behavior
* Aim: Predict future events based on cow behavior pattern
* Solution
    + Understand combined data consisting of cow behavior and events as training dataset for a machine learning problem
    + Use statistical learning model to predict future events
    
    
## Machine/Statistical Learning

* Two categories that are important for us
    1. Supervised 
    2. Unsupervised
    

## Unsupervised Learning

* In unsupervised learning, there is no response, but there is just a set of variables, between which relationships are to be uncovered
* Methods used in unsupervised learning are 
    + Principal Components Analysis (PCA)
    + Cluster analysis
    + Neural Networks


## Supervised Learning

* In supervised learning
    + statistical model is used to understand the relationship between a number of predictor variables (input) and a response variable (output)
    + statistical model can be used to predict future values of the response given new predictor variables
* Methods are 
    + Regression analysis
    + Linear Discriminant analysis
    + Tree-Based Methods (Random Forest, ...)
    + Support Vector Machine (SVM)


## Statistical Learning Application in Everyday Life

* Face recognition based on pictures
* Driving assistance system
* Autonomous driving


## Statistical Learning Application in Agriculture

* Mobile applications for recognizing plants: e.g. PictureThis 
* Weed recognition robots


## Second Level Algorithm in SESAM

* Use overlay dataset of cow behavior and events as training sets to predict the relationship between behavior and the events
* Given new behavior data of cows, new events can be predicted
* Based on predicted events, alerts can be defined


## Details of Second Level Algorithm

* Take an event of interest (heat, calving, mastitis) on a given day
* Separate the cows in two groups
    + 1. cows for which event can be observed
    + 2. cows for which event cannot be observed
* Check differences between cows of two groups in behavior pattern
* Ideally behavior patterns show differences between cows of two groups


## Results of Second Level Algorithm

* How can differences in behavior patterns be found in a systematic way
* Use Support Vector Machine to find optimal border between behavior patterns of two groups
* Use border to predict future occurrences of events based on observed behavior pattern


## Conclusions

* Prototype system of inference of phenotypic events based on behavior data works
* Data quality and density of collected phenotypic events is critical
* Time resolution for different events must be adjusted
* Further research is needed to address open questions


