---
title: "Inferring Cow Phenotypes Based on Sensor Data"
author: "Peter von Rohr"
date: '`r Sys.Date()`'
output:
  powerpoint_presentation: default
  beamer_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Content

1. What are the data 
2. What can we do with this data
3. Introduction to Machine Learning
4. Everyday use cases of machine learning
5. Machine learning in agriculture
6. Second level algorithm
7. Conclusions


## Data

* Output of first level classification
* For every cow over time the basic behavior
* Basic behavior can be one of 
    + Eating
    + Laying
    + Standing
    + Walking
    + Ruminating


## Basic Behavior over Time

```{r tile-plot, echo=FALSE, message=FALSE, warning=FALSE}
n_nr_obs <- 30
vec_behave <- c("eating", "laying", "ruminating", "standing", "walking")
n_nr_behave <- length(vec_behave)
tbl_behavior <- tibble::tibble(x = 1:n_nr_obs, y = sample(vec_behave, n_nr_obs, replace = TRUE))
require(ggplot2)
p <- ggplot(tbl_behavior, aes(x=x, y=as.factor(y), fill = as.factor(y))) + xlab(NULL) + ylab(NULL)
p + geom_tile()
```


## Data Aggregation Over Time

* Summarise duration of behaviors over one day
* Yields Average time for a certain behavior per day
* Defines graphical representation of behavior pattern diagram

<!-- Insert geom_line plot of behavior, similarly to Alibaba or Barplot -->


## Overlay Phenotypes of Interest

* Phenotypes of interest
    + Heat
    + Lameness
    + Calving
    + Mastitis
    + Ketosis
    + ...
* Pooled as events where farmer must be alerted
* Overlay events in behavior pattern diagrams
* Associate events with behavior pattern diagrams


## Aim: Prediction of Events

* Given association of events and cow behavior
* Aim: Predict future events based on cow behavior pattern
* Solution
    + Understand combined data consisting of cow behavior and events as training dataset for a machine learning problem
    + Use statistical learning model to predict future events
    
    
## Machine/Statistical Learning

* Two categories that are important for us
    1. Supervised 
    2. Unsupervised
    
* In supervised learning
    + statistical model is used to understand the relationship between a number of predictor variables (input) and a response variable (output)
    + statistical model can be used to predict future values of the response given new predictor variables
* In unsupervised learning, there is no response, but there is just a set of variables, between which relationships are to be uncovered


## Statistical Learning Application in Everyday Life

* Face recognition based on pictures
* Driving assistence system
* Autonomous driving


## Statistical Learning Application in Agriculture

* Mobile applications for recognising plants: e.g. PictureThis 
* Weed recognition robots


## Second Level Algorithm in SESAM

* Use overlayed dataset of cow behavior and events as training sets to predict the relationship between behavior and the events
* Given new behavior data of cows, new events can be predicted
* Based on predicted events, alerts can be defined


## Conclusions

* Prototype system of inference of phenotypic events based on behavior data works
* Data quality and density of collected phenotypic events is critical
* Time resolution for different events must be adjusted
* Further research is needed to address open questions


